Relative parabolic factorizations of Chevalley groups

### 2015-09-05 paper.tex
Minor fixes in the introduction

\declaretheorem[name=Theorem, Refname={Theorem,Theorems}]{thm}
\declaretheorem[name=Corollary, Refname={Corollary, Corollaries}, sibling=thm]{cor}

\declaretheorem[name=Lemma, Refname={Lemma,Lemmas}]{lemma}
\declaretheorem[name=Proposition, Refname={Proposition,Propositions}, sibling=lemma]{prop}

@@

Our first application of parabolic factorizations is the following theorem which gives an estimate of the width of $\G(\Phi, R, I)$ in terms of certain subsets of $\mathcal{Z}$.
\begin{thm}\label{thm:width} Let $\Phi$ be an irreducible root system, and let $I$ be an ideal of a commutative ring $R$.
\begin{enumerate}
\item If $\sr(I) = 1$, then $W(\E(\Phi,R,I),\mathcal{Z}(-\Pi))\leqslant 3|\Phi^+|+2\rk(\Phi)-1$;
\item Let $p$ be a prime integer, set $R = \mathbb{Z}[\sfrac{1}{p}]$ and assume $\Phi$ is classical. Then under the assumption of the Generalized Riemann Hypothesis one has
%TODO: Now check the coefficients of \rk(\Phi) in the expression below
\begin{alignat*}{2}
& W(\E(\Phi,R,I), \mathcal{Z}(\Sigma_\ell^-))\leqslant 3|\Phi^+| + 2\rk(\Phi) + 1 & \text{ if } \Phi=\rA_\ell,\rC_\ell, \\
& W(\E(\Phi,R,I), \mathcal{Z}(\Sigma_\ell^-))\leqslant 4|\Phi^+| + \rk(\Phi) + 1 & \text{ if } \Phi=\rB_\ell,\rD_\ell;
\end{alignat*}
\item Let $\mathcal{O}_S$ be a Dedekind ring of arithmetic type embedded into some global field $k$ having a real embedding. 
Assume $\Phi$ is classical of rank $\geqslant2$, then $W(\G(\Phi,\mathcal{O}_S,I), \mathcal{Z}(\Sigma_\ell^-))$ is finite.
\end{enumerate}
\end{thm}
The particular case $\Phi = \rA_\ell$, $R = I = \mathbb{Z}$ of the above theorem is similar in spirit to a recent theorem of U.~Hadad (see ~\cite[Theorem~1.6]{Ha12}).

@@

It is also classically known that over an arbitrary field $F$ the group $\SL(n+1, F)$, $n\geq 3$ is a product of at most $4$ subgroups of type $\SL(n, F)$.
On the other hand, from the classical Dennis---Vaserstein theorem (see~\cite{Va69} and \cite[Lemma~2.1]{ST76})
it follows that $\SL(n+1, R, I)$ is a product of a most $5$ subgroups of type $\SL_n(R, I)$ for any ideal $I$ satisfying $\sr(I)\leqslant n-1$.
Using another Dennis---Vaserstein type factorization we obtain the following result.
\begin{thm}\label{thm:spin-sln-prod}
Assume that $\sr(I) \leqslant 2$ then the group $\Epin_{2\ell}(R, I)=\E_{sc}(\rD_\ell, R, I)$ is a product of at most $9$ copies of subgroups of type $\rA_{\ell-1}$.
\end{thm}

>>>>

\declaretheorem[name=Theorem, Refname={Theorem,Theorems}]{thm}
\declaretheorem[name=Proposition, Refname={Proposition,Propositions}, sibling=thm]{prop}
\declaretheorem[name=Corollary, Refname={Corollary, Corollaries}, numberwithin=section]{cor}
\declaretheorem[name=Lemma, Refname={Lemma,Lemmas}, numberwithin=section]{lemma}

@@

Our first application of parabolic factorizations is the following theorem which gives an estimate of the width of $\G(\Phi, R, I)$ in terms of certain subsets of $\mathcal{Z}$.
\begin{thm}\label{thm:width} Let $\Phi$ be an irreducible root system, and let $I$ be an ideal of a commutative ring $R$.
\begin{enumerate}
\item If $\sr(I) = 1$, then $W(\E(\Phi,R,I),\mathcal{Z}(\Pi))\leqslant 3|\Phi^+|+2\rk(\Phi)-1$;
\item Let $p$ be a prime integer, set $R = \mathbb{Z}[\sfrac{1}{p}]$ and assume $\Phi$ is classical. Then under the assumption of the Generalized Riemann Hypothesis one has
%TODO: Now check the coefficients of \rk(\Phi) in the expression below
\begin{alignat*}{2}
& W(\E(\Phi,R,I), \mathcal{Z}(\Sigma_\ell^-))\leqslant 3|\Phi^+| + 2\rk(\Phi) + 1 & \text{ if } \Phi=\rA_\ell,\rC_\ell, \\
& W(\E(\Phi,R,I), \mathcal{Z}(\Sigma_\ell^-))\leqslant 4|\Phi^+| + \rk(\Phi) + 1 & \text{ if } \Phi=\rB_\ell,\rD_\ell;
\end{alignat*}
\item Let $\mathcal{O}_S$ be a Dedekind ring of arithmetic type embedded into some global field $k$ having a real embedding. 
Assume $\Phi$ is classical of rank $\geqslant2$, then $W(\G(\Phi,\mathcal{O}_S,I), \mathcal{Z}(\Sigma_\ell^-))$ is finite.
\end{enumerate}
\end{thm}
The particular case $\Phi = \rA_\ell$, $R = I = \mathbb{Z}$ of the above theorem is similar in spirit to a recent theorem of U.~Hadad (see ~\cite[Theorem~1.6]{Ha12}).

@@

It is also classically known that over an arbitrary field $F$ the group $\SL_{n+1}(F)$, $n\geq 3$ is a product of at most $4$ subgroups of type $\SL(n, F)$.
On the other hand, from the classical Dennis---Vaserstein theorem (see~\cite{Va69} and \cite[Lemma~2.1]{ST76})
it follows that $\SL_{n+1}(R, I)$ is a product of a most $5$ subgroups of type $\SL_n(R, I)$ for any ideal $I$ satisfying $\sr(I)\leqslant n-1$.
Using another Dennis---Vaserstein type factorization we obtain the following result.
\begin{thm}\label{thm:spin-sln-prod}
Assume that $\sr(I) \leqslant 2$ then the group $\Epin_{2\ell}(R, I)=\E(\rD_\ell, R, I)$ is a product of at most $9$ copies of its subgroups of type $\rA_{\ell-1}$.
\end{thm}
